---
title: "Collate similarities"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    toc_depth: 3
    number_sections: yes
    theme: lumen
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
params:
  batch: "2020_11_04_CPJUMP1"
  experiment: !r data.frame(Metadata_cell_line = "U2OS", Metadata_experiment_condition = "Standard")
  all_same_cols: !r NULL
  normalization: whole_plate
  debug_all_same_cols: !r c("Metadata_experiment_type")
  debug_experiment: !r data.frame(Metadata_cell_line = "U2OS", Metadata_timepoint = "48", Metadata_experiment_type = "Compound", Metadata_experiment_condition = "Standard")
---

# Setup

```{r message=FALSE}
library(magrittr)
library(tidyverse)
library(glue)
library(arrow)
```


```{r}
source("https://raw.githubusercontent.com/shntnu/grit-benchmark/47ae58e83e3db85eeaf43b51cfe0dc5a187410cb/R/cytominer-eval.R")
```


```{r}
experiment <- params$experiment

if(!is.null(experiment)) {
  experiment_tag <-
    experiment %>%
    unite("experiment_tag", everything()) %>%
    pull("experiment_tag") %>%
    sort() %>%
    paste(collapse = "_")
} else {
  experiment_tag <- "all"
}

all_same_cols <- params$all_same_cols

if(!is.null(all_same_cols)) {
  all_same_cols_tag <-
    all_same_cols %>%
    str_remove_all("Metadata_") %>%
    sort() %>%
    paste(collapse = "_")
} else {
  all_same_cols_tag <- "none"
}

all_same_cols_tag <- paste0("same_", all_same_cols_tag)

comparision_tag <- paste(experiment_tag, all_same_cols_tag, sep = "_")

futile.logger::flog.info(glue("Comparison tag = {comparision_tag}"))
```


```{r}
if (params$normalization == "whole_plate") {
  norm_suffix <- ""
} else if (params$normalization == "negcon") {
  norm_suffix <- "_negcon"
}
sprintf("Normalization = %s. Using suffix = '%s'", params$normalization, norm_suffix)
```


```{r}
batch <- params$batch
futile.logger::flog.info(glue("Batch = {batch}"))
```


```{r}
filename_prefix <- glue("{batch}_{comparision_tag}_normalized_feature_select_outlier_trimmed{norm_suffix}")
```


# Load similarity matrix on Level 4b

```{r}
futile.logger::flog.info("Loading similarity ...")

parquet_file <- 
  glue("{batch}/{filename_prefix}_sim.parquet")

sim_df <-
    sim_read(parquet_file)
```

# Compute similarity sets

## Filter out some rows

```{r}
drop_group <-
  NULL
```

## Similarity to reference

Fetch similarities between

a. all rows (except, optionally those containing `reference`)

and 

b. all rows containing `reference`

Do so only for those (a, b) pairs that 

- have *same* values in *all* columns of `all_same_cols_ref`

```{r}
reference <-
  data.frame(Metadata_negcon_or_other = "negcon")

all_same_cols_ref <-
  c("Metadata_cell_line",
    "Metadata_timepoint",
    "Metadata_experiment_condition",
    "Metadata_experiment_type",
    "Metadata_Plate")
```

## Similarity to replicates (no references)

Fetch similarities between

a. all rows except `reference` rows

and

b. all rows except `reference` rows (i.e. to each other)

Do so for only those (a, b) pairs that 

- have *same* values in *all* columns of `all_same_cols_rep`

Keep, both, (a, b) and (b, a)

```{r}
all_same_cols_rep <-
  c("Metadata_cell_line",
    "Metadata_timepoint",
    "Metadata_experiment_condition",
    "Metadata_experiment_type",
    "Metadata_broad_sample",
    "Metadata_pert_iname",
    "Metadata_negcon_or_other",
    "Metadata_Well") # including well here
```

## Similarity to replicates (only references)

Fetch similarities between

a. all rows containing `reference`

and

b. all rows containing `reference` (i.e. to each other)

Do so for only those (a, b) pairs that 

- have *same* values in *all* columns of `all_same_cols_rep_ref`. 

Keep, both, (a, b) and (b, a)

```{r}
all_same_cols_rep_ref <-
  c(
    "Metadata_cell_line",
    "Metadata_timepoint",
    "Metadata_experiment_condition",
    "Metadata_experiment_type",
    "Metadata_broad_sample",
    "Metadata_pert_iname",
    "Metadata_negcon_or_other",
    "Metadata_Well"
  )
```

## Similarity to non-replicates

Fetch similarities between

a. all rows (except, optionally, `reference` rows)

and

b. all rows except `reference` rows

Do so for only those (a, b) pairs that 

- have *same* values in *all* columns of `all_same_cols_non_rep`

- have *different* values in *all* columns `all_different_cols_non_rep`

- have *different* values in *at least one* column of `any_different_cols_non_rep`

Keep, both, (a, b) and (b, a)

```{r}
any_different_cols_non_rep <-
  c("Metadata_cell_line",
    "Metadata_broad_sample",
    "Metadata_timepoint",
    "Metadata_Well")
all_same_cols_non_rep <-
  c("Metadata_cell_line",
    "Metadata_experiment_condition",
    "Metadata_experiment_type",
    "Metadata_timepoint",
    "Metadata_Plate")
all_different_cols_non_rep <-
  c("Metadata_target")
```

## Similarity to group

Fetch similarities between

a. all rows 

and 

b. all rows

Do so only for those (a, b) pairs that 

- have *same* values in *all* columns of `all_same_cols_group`

- have *different* values in *at least one* column of `any_different_cols_group`

```{r}
all_same_cols_group <-
  c("Metadata_cell_line",
    "Metadata_experiment_condition",
    "Metadata_experiment_type",
    "Metadata_timepoint",
    "Metadata_target")
any_different_cols_group <-
  c("Metadata_cell_line",
    "Metadata_timepoint",
    "Metadata_broad_sample")
```

## Combine all and annotate the similarity matrix

```{r}
annotation_cols <-
  c("Metadata_cell_line",
    "Metadata_timepoint",
    "Metadata_experiment_condition",
    "Metadata_experiment_type",
    "Metadata_broad_sample",
    "Metadata_pert_iname",
    "Metadata_negcon_or_other",
    "Metadata_Well")

collated_sim <-
  sim_collate(
    sim_df,
    reference,
    all_same_cols_rep = all_same_cols_rep,
    all_same_cols_rep_ref = all_same_cols_rep_ref,
    all_same_cols_ref = all_same_cols_ref,
    any_different_cols_non_rep = any_different_cols_non_rep,
    all_same_cols_non_rep = all_same_cols_non_rep,
    all_different_cols_non_rep = all_different_cols_non_rep,
    any_different_cols_group = any_different_cols_group,
    all_same_cols_group = all_same_cols_group,
    annotation_cols = annotation_cols,
    drop_group = drop_group
  )
```

# Compute metrics

```{r}
norm_non_rep_metrics <- 
  sim_metrics(collated_sim, "non_rep", calculate_grouped = TRUE)

norm_ref_metrics <- 
  sim_metrics(collated_sim, "ref", calculate_grouped = TRUE)

level_1_0_metrics <-
  inner_join(
    norm_non_rep_metrics[["per_row"]],
    norm_ref_metrics[["per_row"]]
  )

level_1_metrics <-
  inner_join(
    norm_non_rep_metrics[["per_set"]],
    norm_ref_metrics[["per_set"]]
  )

level_2_1_metrics <-
  inner_join(
    norm_non_rep_metrics[["per_set_group"]],
    norm_ref_metrics[["per_set_group"]]
  )

level_1_and_2_1_metrics <-
  level_1_metrics %>%
  inner_join(level_2_1_metrics, by = c(all_same_cols_rep))


metric_set_names <- c(
  "level_1_0_metrics",
  "level_1_metrics",
  "level_2_1_metrics",
  "level_1_and_2_1_metrics"
)

walk(metric_set_names, function(metric_set_name) {
  parquet_file <-
    glue("{batch}/{filename_prefix}_{metric_set_name}.parquet")
  
  futile.logger::flog.info(glue("Writing {parquet_file} ..."))
  
  eval(sym(metric_set_name)) %>%
    arrow::write_parquet(glue(parquet_file), compression = "gzip", compression_level = 9)
  
})

parquet_file <-
  glue("{batch}/{filename_prefix}_sim_collated.parquet")

futile.logger::flog.info(glue("Writing {parquet_file} ..."))

collated_sim %>%
  arrow::write_parquet(glue(parquet_file), compression = "gzip", compression_level = 9)
```

