---
title: "Measure similarities"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    toc_depth: 3
    number_sections: yes
    theme: lumen
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
params:
  batch: "2020_11_04_CPJUMP1"
  experiment: 
    value:
      Metadata_experiment_condition: Standard
  strata:
    value:
      - Metadata_experiment_condition
      - Metadata_experiment_type
      - Metadata_cell_line
      - Metadata_timepoint
  data_level: normalized_feature_select_outlier_trimmed
  normalization: whole_plate
  similarity_method: cosine
---

# Setup

```{r message=FALSE}
library(magrittr)
library(tidyverse)
library(glue)
library(arrow)
library(matric)
library(logger)
source("utils.R")
log_layout(layout_no_timestamp)
```


```{r}
process_notebook_params(params)

log_info("Comparison tag = {comparision_tag}")

log_info("Normalization = {normalization}. Using suffix = '{norm_suffix}'")

log_info("Batch = {batch}")

log_info("Data level = {data_level}")

log_info("Data level tag = {data_level_tag}")

log_info("Similarity method = {similarity_method}")

log_info("Filename prefix = {filename_prefix}")
```

# Load profiles

```{r}
filename_prefix_profiles <- glue("{batch}_all_{data_level}{norm_suffix}")

parquet_file <- glue("../collated/{batch}/{filename_prefix_profiles}.parquet")

log_info(glue("Loading {parquet_file} ..."))

stopifnot(file.exists(parquet_file))

profiles <-
  read_parquet(parquet_file) 

if(!is.null(experiment))  {
  profiles %<>% inner_join(experiment)
  
}
```

Create a new column `Metadata_pert_iname_aug` so that controls and empty wells in different positions are unique perturbations.
This is better than using `Metadata_Well` as an additional grouping columns because we don't want treatment wells to be handled this way (entirely because ORFs have 2 replicates on the same plate)

```{r}
profiles <-
  profiles %>%
  mutate(Metadata_pert_iname_aug = ifelse(
    Metadata_pert_type == "trt",
    Metadata_pert_iname,
    paste(Metadata_pert_iname, Metadata_Well, sep = "_")
  ))
```


```{r}
profiles %>%
  distinct(Metadata_experiment_condition, Metadata_experiment_type, Metadata_Plate) %>%
  group_by(Metadata_experiment_condition, Metadata_experiment_type) %>%
  tally()
```

# Inspect negcon annotations

```{r}
profiles %>%
  group_by(Metadata_experiment_condition, Metadata_experiment_type, Metadata_negcon_or_other, Metadata_negcon_control_type_trimmed) %>%
  tally()
```


```{r}
profiles %>%
  filter(Metadata_experiment_type != "Compound") %>%
  filter(is.na(Metadata_pert_iname) | is.na(Metadata_target)) %>%
  group_by(Metadata_experiment_condition, Metadata_experiment_type, Metadata_negcon_or_other, Metadata_pert_iname, Metadata_target) %>%
  tally()
```


# Compute similarity matrix

```{r}
log_info("Calculating similarity ...")

similarity_method <- params$similarity_method

sim_df <-
  matric::sim_calculate(population = profiles,
                        strata = strata, 
                        method = similarity_method)
parquet_file <- 
  glue("{batch}/{filename_prefix}_sim_{similarity_method}.parquet")

sim_df %>%
  sim_write(parquet_file)
```


