---
title: "Collate and reprocess CPJUMP1 profiles"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    toc_depth: 3
    number_sections: yes
    theme: lumen
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
params:
  batch: 2020_11_04_CPJUMP1
---

This notebook does 3 things

1. Produces collated versions of the `2020_11_04_CPJUMP1` Level 3 (`_augmented`) and Level 4a (`_normalized`) as [Parquet](https://parquet.apache.org/) files.
2. Recomputes Level 4b (`_normalized_feature_select`) on the collated Level 4a file.
3. Computes another version of the Level 4b file (`normalized_feature_select_outlier_trimmed`) files with the outliers removed. using bespoke logic for this dataset. This should be done directly in the profiling recipe in the future, once we have figured out how to handle outliers better.

The notebook generates several intermediate parquet files for convenience, but they are not committed to the repository.

The notebook also requires executing a `pycytominer` script by hand (hacky! but reproducible)

```
R version 4.0.3 (2020-10-10)
Platform: x86_64-apple-darwin17.0 (64-bit)
Running under: macOS Catalina 10.15.7

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
LAPACK: /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRlapack.dylib

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] arrow_3.0.0     glue_1.4.2      forcats_0.5.0   stringr_1.4.0   dplyr_1.0.4     purrr_0.3.4     readr_1.4.0     tidyr_1.1.2    
 [9] tibble_3.0.6    ggplot2_3.3.3   tidyverse_1.3.0 magrittr_2.0.1 

loaded via a namespace (and not attached):
 [1] Rcpp_1.0.6           cellranger_1.1.0     pillar_1.4.7         compiler_4.0.3       RColorBrewer_1.1-2   formatR_1.7         
 [7] futile.logger_1.4.3  dbplyr_2.1.0         futile.options_1.0.1 tools_4.0.3          bit_4.0.4            lubridate_1.7.9.2   
[13] jsonlite_1.7.2       lifecycle_1.0.0      gtable_0.3.0         pkgconfig_2.0.3      rlang_0.4.10         reprex_0.3.0        
[19] cli_2.3.0            rstudioapi_0.13      DBI_1.1.1            haven_2.3.1          xfun_0.21            withr_2.4.1         
[25] xml2_1.3.2           httr_1.4.2           knitr_1.31           fs_1.5.0             generics_0.1.0       vctrs_0.3.6         
[31] platetools_0.1.3     hms_1.0.0            bit64_4.0.5          grid_4.0.3           tidyselect_1.1.0     R6_2.5.0            
[37] readxl_1.3.1         modelr_0.1.8         lambda.r_1.2.4       scales_1.1.1         backports_1.2.1      ellipsis_0.3.1      
[43] rvest_0.3.6          assertthat_0.2.1     colorspace_2.0-0     stringi_1.5.3        munsell_0.5.0        broom_0.7.5         
[49] crayon_1.4.1   
```

# Setup

```{r message=FALSE}
library(magrittr)
library(tidyverse)
library(glue)
library(arrow)
```

```{r}
experiment_filter <- NULL
```

```{r}
batch <- params$batch
```

```{r}
overwrite_level_3_profiles <- FALSE
overwrite_level_4a_profiles <- FALSE
overwrite_level_4b_profiles <- FALSE
```

```{r}
mapping <-
  tribble(
    ~Metadata_Plate, ~Metadata_cell_line, ~Metadata_timepoint, ~Metadata_experiment_type, ~Metadata_experiment_condition, ~Metadata_replicate_id,
    "BR00118050", "A549", "96", "ORF", "Standard", 1,
    "BR00117006", "A549", "96", "ORF", "Standard", 2,
    "BR00118039", "U2OS", "96", "ORF", "Standard", 1,
    "BR00118040", "U2OS", "96", "ORF", "Standard", 2,
    "BR00117020", "A549", "48", "ORF", "Standard", 1,
    "BR00117021", "A549", "48", "ORF", "Standard", 2,
    "BR00117022", "U2OS", "48", "ORF", "Standard", 1,
    "BR00117023", "U2OS", "48", "ORF", "Standard", 2,
    "BR00118041", "A549", "96", "CRISPR", "Standard", 1,
    "BR00118042", "A549", "96", "CRISPR", "Standard", 2,
    "BR00118043", "A549", "96", "CRISPR", "Standard", 3,
    "BR00118044", "A549", "96", "CRISPR", "Standard", 4,
    "BR00118045", "U2OS", "96", "CRISPR", "Standard", 1,
    "BR00118046", "U2OS", "96", "CRISPR", "Standard", 2,
    "BR00118047", "U2OS", "96", "CRISPR", "Standard", 3,
    "BR00118048", "U2OS", "96", "CRISPR", "Standard", 4,
    "BR00117003", "A549", "144", "CRISPR", "Standard", 1,
    "BR00117004", "A549", "144", "CRISPR", "Standard", 2,
    "BR00117005", "A549", "144", "CRISPR", "Standard", 3,
    "BR00117000", "A549", "144", "CRISPR", "Standard", 4,
    "BR00116997", "U2OS", "144", "CRISPR", "Standard", 1,
    "BR00116998", "U2OS", "144", "CRISPR", "Standard", 2,
    "BR00116999", "U2OS", "144", "CRISPR", "Standard", 3,
    "BR00116996", "U2OS", "144", "CRISPR", "Standard", 4,
    "BR00116991", "A549", "24", "Compound", "Standard", 1,
    "BR00116992", "A549", "24", "Compound", "Standard", 2,
    "BR00116993", "A549", "24", "Compound", "Standard", 3,
    "BR00116994", "A549", "24", "Compound", "Standard", 4,
    "BR00116995", "U2OS", "24", "Compound", "Standard", 1,
    "BR00117024", "U2OS", "24", "Compound", "Standard", 2,
    "BR00117025", "U2OS", "24", "Compound", "Standard", 3,
    "BR00117026", "U2OS", "24", "Compound", "Standard", 4,
    "BR00117017", "A549", "48", "Compound", "Standard", 1,
    "BR00117019", "A549", "48", "Compound", "Standard", 2,
    "BR00117015", "A549", "48", "Compound", "Standard", 3,
    "BR00117016", "A549", "48", "Compound", "Standard", 4,
    "BR00117012", "U2OS", "48", "Compound", "Standard", 1,
    "BR00117013", "U2OS", "48", "Compound", "Standard", 2,
    "BR00117010", "U2OS", "48", "Compound", "Standard", 3,
    "BR00117011", "U2OS", "48", "Compound", "Standard", 4,
    "BR00118049", "A549", "96", "ORF", "Blasticidin", 1,
    "BR00117002", "A549", "144", "CRISPR", "Puromycin", 1,
    "BR00117001", "A549", "144", "CRISPR", "Puromycin", 2,
    "BR00117054", "A549", "48", "Compound", "p20pc_Seed_Density", 1,
    "BR00117055", "A549", "48", "Compound", "p20pc_Seed_Density", 2,
    "BR00117008", "A549", "48", "Compound", "m20pc_Seed_Density", 1,
    "BR00117009", "A549", "48", "Compound", "m20pc_Seed_Density", 2,
    "BR00117052", "A549", "48", "Compound", "Cas9", 1,
    "BR00117053", "A549", "48", "Compound", "Cas9", 2,
    "BR00117050", "A549", "48", "Compound", "Cas9", 3,
    "BR00117051", "A549", "48", "Compound", "Cas9", 4
  )
```

Read additional metadata

```{r}

basedir <- "https://raw.githubusercontent.com/jump-cellpainting/JUMP-Target/40838ab860ee9b872a9993105d5b97d6bef2dab5/"

compound_metadata <-
  read_tsv(
    glue("{basedir}/JUMP-Target_compound_metadata.tsv"),
    col_types = cols_only(broad_sample = col_character(),
                          pert_iname = col_character(),
                          pubchem_cid = col_character(),
                          InChIKey = col_character())
  ) %>%
  select(Metadata_broad_sample = broad_sample,
         Metadata_pert_iname = pert_iname,
         Metadata_pubchem_cid = pubchem_cid,
         Metadata_InChIKey = InChIKey)

crispr_metadata <-
  read_tsv(
    glue("{basedir}/JUMP-Target_crispr_metadata.tsv"),
    col_types = cols_only(broad_sample = col_character(),
                          negcon_control_type = col_character(),
                          gene = col_character(),
                          target_sequence = col_character())
  ) %>%
  select(Metadata_broad_sample = broad_sample,
         Metadata_negcon_control_type = negcon_control_type,
         Metadata_gene = gene,
         Metadata_target_sequence = target_sequence)

orf_metadata <-
  read_tsv(
    glue("{basedir}/JUMP-Target_orf_metadata.tsv"),
    col_types = cols_only(broad_sample = col_character(),
                          gene = col_character())
  ) %>%
  select(Metadata_broad_sample = broad_sample,
         Metadata_gene = gene)

additional_metadata <-
  list(ORF = orf_metadata,
       CRISPR = crispr_metadata,
       Compound = compound_metadata)

rm(list = c("orf_metadata", "crispr_metadata", "compound_metadata"))
```

coltypes map

```{r}
col_types_map <-
  list(
    ORF = cols(
      .default = col_double(),
      Metadata_plate_map_name = col_character(),
      Metadata_broad_sample = col_character(),
      Metadata_Plate = col_character(),
      Metadata_Well = col_character(),
      Metadata_genes = col_character(),
      Metadata_pert_type = col_character(),
      Metadata_control_type = col_character(),
      Metadata_Assay_Plate_Barcode = col_character(),
      Metadata_Plate_Map_Name = col_character()
    ),
    CRISPR = cols(
      .default = col_double(),
      Metadata_plate_map_name = col_character(),
      Metadata_broad_sample = col_character(),
      Metadata_Plate = col_character(),
      Metadata_Well = col_character(),
      Metadata_genes = col_character(),
      Metadata_pert_type = col_character(),
      Metadata_control_type = col_character(),
      Metadata_Assay_Plate_Barcode = col_character(),
      Metadata_Plate_Map_Name = col_character()
    ),
    Compound = cols(
      .default = col_double(),
      Metadata_plate_map_name = col_character(),
      Metadata_broad_sample = col_character(),
      Metadata_mg_per_ml = col_double(),
      Metadata_mmoles_per_liter = col_double(),
      Metadata_solvent = col_character(),
      Metadata_Plate = col_character(),
      Metadata_Well = col_character(),
      Metadata_target = col_character(),
      Metadata_pert_type = col_character(),
      Metadata_control_type = col_character(),
      Metadata_Assay_Plate_Barcode = col_character(),
      Metadata_Plate_Map_Name = col_character()
    )
  )

col_types_master <- cols(
  .default = col_double(),
  Metadata_plate_map_name = col_character(),
  Metadata_broad_sample = col_character(),
  Metadata_Plate = col_character(),
  Metadata_Well = col_character(),
  Metadata_genes = col_character(),
  Metadata_pert_type = col_character(),
  Metadata_control_type = col_character(),
  Metadata_Assay_Plate_Barcode = col_character(),
  Metadata_Plate_Map_Name = col_character(),
  Metadata_cell_line = col_character(),
  Metadata_timepoint = col_character(),
  Metadata_experiment_type = col_character(),
  Metadata_experiment_condition = col_character(),
  Metadata_replicate_id = col_character(),
  Metadata_gene = col_character(),
  Metadata_negcon_control_type = col_character(),
  Metadata_target_sequence = col_character(),
  Metadata_mg_per_ml = col_double(),
  Metadata_mmoles_per_liter = col_double(),
  Metadata_solvent = col_character(),
  Metadata_target = col_character(),
  Metadata_pert_or_negcon = col_character(),
  Metadata_pert_iname = col_character(),
  Metadata_pubchem_cid = col_character(),
  Metadata_InChIKey = col_character()
)
```


```{r}
experiment_spec_cols <- c(
  "Metadata_cell_line",
  "Metadata_timepoint",
  "Metadata_experiment_type",
  "Metadata_experiment_condition"
)

experiments <-
  mapping %>%
  distinct(across(all_of(experiment_spec_cols))) %>%
  unite("Metadata_experiment_tag", everything(), remove = FALSE)
```


```{r}
if(!is.null(experiment_filter)) {
  experiments %<>% inner_join(experiment_filter)
  
  experiment_filter_tag <- paste(experiment_filter, collapse = "_")
  
  experiment_filter_tag <-
    experiment_filter %>%
    unite("experiment_filter_tag", everything()) %>%
    pull("experiment_filter_tag") %>%
    sort() %>%
    paste(collapse = "_")
} else {
  experiment_filter_tag = "all"
}
futile.logger::flog.info(glue("Experiment tag = {experiment_filter_tag}")
)
```


```{r}
report_na_stats <- function(df) {
  variables <- str_subset(colnames(df), "^Metadata_", negate = T)
  
  df %>%
    summarise(across(all_of(variables), ~ any(is.na(.)))) %>%
    pivot_longer(everything(), values_to = "is_na_feature") %>%
    group_by(is_na_feature) %>%
    tally() %>%
    ungroup() %>%
    pivot_wider(
      names_from = "is_na_feature",
      values_from = "n",
      names_prefix = "is_na_"
    ) %>%
    print()
}
```


```{r}
find_outlier_variables <-
  function(df,
           outlier_variable_value_threshold = 20,
           outlier_well_frac_threshold_all_plates = .05,
           outlier_well_frac_threshold_per_plate = .05,
           outlier_plate_frac_threshold = 0.05,
           n_wells_per_plate = 384) {
    df %<>% ungroup()
    
    variables <- str_subset(colnames(df), "^Metadata_", negate = T)
    
    n_plates <- df %>% distinct(Metadata_Plate) %>% nrow()
    n_variables <- length(variables)
    n_wells <- nrow(df)
    
    # function to flag a value as an outlier
    outlier_well_count <- function(x)
      sum(abs(x) > outlier_variable_value_threshold)
    
    # for each variable, count the number of outlier wells per plate
    outlier_df <-
      df %>%
      group_by(Metadata_Plate) %>%
      summarise(across(
        all_of(variables),
        list(outlier_well_count = outlier_well_count)
      )) %>%
      pivot_longer(-Metadata_Plate,
                   values_to = "outlier_well_count",
                   names_to = "variable_name")
    
    # variables that are outliers in more than `p` fraction of wells across all plates
    # where `p` = outlier_well_frac_threshold_all_plates
    outlier_variables_stat_well <-
      outlier_df %>%
      group_by(variable_name) %>%
      summarize(outlier_well_frac = sum(outlier_well_count) / n_wells) %>%
      filter(outlier_well_frac > outlier_well_frac_threshold_all_plates)
    
    # variables that are outliers in more than `p` fraction of plates
    # a variable is an outlier in a plate if it is an outlier in more than `q`
    # fraction of wells,
    # where `p` = outlier_well_frac_threshold_all_plates
    # where `q` = outlier_well_frac_threshold_per_plate
    outlier_variables_stat_plate <-
      outlier_df %>%
      filter(outlier_well_count >
               n_wells_per_plate *
               outlier_well_frac_threshold_per_plate) %>%
      group_by(variable_name) %>%
      count(name = "outlier_plate_count") %>%
      mutate(outlier_plate_frac = outlier_plate_count / n_plates) %>%
      filter(outlier_plate_frac > outlier_plate_frac_threshold)
    
    outlier_variables <-
      c(
        outlier_variables_stat_well %>%
          pull(variable_name),
        outlier_variables_stat_plate %>%
          pull(variable_name)
      ) %>%
      str_remove("_outlier_well_count")
    
    outlier_variables
  }
```


```{r}
collate_profiles <-
  function(experiments,
           batch,
           profile_level,
           mapping,
           col_types_map,
           additional_metadata,
           experiment_spec_cols,
           report_stats = TRUE,
           overwrite_profiles = FALSE,
           write_csvgz = FALSE,
           load_if_present = FALSE,
           experiment_filter_tag = "all",
           additional_metadata_join_cols =
             c("Metadata_broad_sample"),
           mapping_join_cols =
             c("Metadata_Plate")) {
    parquet_all_experiments_file <-
      glue("../collated/{batch}/{batch}_{experiment_filter_tag}_{profile_level}.parquet")
    
    csvgz_all_experiments_file <-
      glue("../collated/{batch}/{batch}_{experiment_filter_tag}_{profile_level}.csv.gz")
    
    if (!file.exists(parquet_all_experiments_file)) {
      for (i in 1:nrow(experiments)) {
        experiment <- experiments[i,]
        
        experiment_tag <- experiment$Metadata_experiment_tag
        
        parquet_file <-
          glue("../collated/{batch}/{batch}_{experiment_tag}_{profile_level}.parquet")
        
        futile.logger::flog.threshold(futile.logger::INFO)
        
        if (!file.exists(parquet_file) | overwrite_profiles) {
          futile.logger::flog.info("Reading CSVs ...")
          
          profiles <-
            mapping %>%
            inner_join(experiment, by = experiment_spec_cols) %>%
            pull("Metadata_Plate") %>%
            map_df(function(plate) {
              filename <-
                glue("../profiles/{batch}/{plate}/{plate}_{profile_level}.csv.gz")
              read_csv(filename, col_types = col_types_map[[experiment$Metadata_experiment_type]])
            })
          
          profiles %<>% inner_join(mapping, by = mapping_join_cols)
          
          if (experiment$Metadata_experiment_type == "Compound") {
            profiles %<>% mutate(Metadata_broad_sample = ifelse(
              is.na(Metadata_broad_sample),
              "DMSO",
              Metadata_broad_sample
            ))
            
            profiles %<>% mutate(
              Metadata_pert_or_negcon =
                ifelse(
                  Metadata_pert_type == "control" &
                    Metadata_control_type == "negcon",
                  "negcon",
                  "pert"
                )
            )
          }
          
          profiles %<>%
            left_join(additional_metadata[[experiment$Metadata_experiment_type]],
                      by = additional_metadata_join_cols) %>%
            replace_na(
              list(
                Metadata_gene = "Unknown",
                Metadata_InChIKey = "Unknown",
                Metadata_negcon_control_type = "Unknown",
                Metadata_pert_iname = "Unknown",
                Metadata_pubchem_cid = "Unknown",
                Metadata_target_sequence = "Unknown"
              )
            )
          
          futile.logger::flog.info(glue("Writing {parquet_file} ..."))
          
          profiles %>%
            select(matches("Metadata_"), everything()) %>%
            write_parquet(parquet_file)
          
        } else {
          futile.logger::flog.info(glue("{parquet_file} already exists."))
        }
      }
      
      if (!file.exists(csvgz_all_experiments_file) ||
          !file.exists(parquet_all_experiments_file) ||
          overwrite_profiles) {
        profiles <-
          experiments %>%
          pull("Metadata_experiment_tag") %>%
          map_df(function(experiment_tag) {
            parquet_file <-
              glue("../collated/{batch}/{batch}_{experiment_tag}_{profile_level}.parquet")
            
            futile.logger::flog.info(glue("Loading {parquet_file} ..."))
            
            read_parquet(parquet_file)
            
          })
        profiles %<>% select(matches("Metadata_"), everything())

      }
      
      if (!file.exists(csvgz_all_experiments_file) ||
          overwrite_profiles) {
        futile.logger::flog.info(glue("Writing {csvgz_all_experiments_file} ..."))
        
        profiles %>%
          write_csv(csvgz_all_experiments_file)
      }
      
      if (!file.exists(parquet_all_experiments_file) ||
          overwrite_profiles) {
      
        futile.logger::flog.info(glue("Writing {parquet_all_experiments_file} ..."))
        
        profiles %>%
          write_parquet(parquet_all_experiments_file, compression = "gzip", compression_level = 9)
        
      }
    }
    
    futile.logger::flog.info(glue("{parquet_all_experiments_file} is available."))
    
    if (load_if_present) {
      futile.logger::flog.info(glue("Loading {parquet_all_experiments_file} ..."))
      
      profiles <-
        read_parquet(parquet_all_experiments_file)
      
      if (report_stats) {
        report_na_stats(profiles)
        
        n_outlier_variables <-
          length(find_outlier_variables(profiles))
        
        futile.logger::flog.info(
          glue(
            "There are {n_outlier_variables} outlier variables assuming they were z-scored."
          )
        )
      }
      
      profiles
      
    } else {
      NULL
      
    }
    
  }
```


# Load and save Level 3 profiles

```{r}
profiles <- 
  collate_profiles(
  experiments = experiments,
  profile_level = "augmented",
  batch = batch,
  overwrite_profiles = FALSE,
  experiment_filter_tag = experiment_filter_tag,
  report_stats = TRUE,
  mapping = mapping,
  col_types_map = col_types_map,
  additional_metadata = additional_metadata,
  experiment_spec_cols = experiment_spec_cols,
  load_if_present = TRUE,
  additional_metadata_join_cols =
    c("Metadata_broad_sample"),
  mapping_join_cols =
    c("Metadata_Plate")
) 

```


# Load and save Level 4a profiles

```{r}
profiles <- 
  collate_profiles(
  experiments = experiments,
  profile_level = "normalized",
  batch = params$batch,
  overwrite_profiles = FALSE,
  write_csvgz = TRUE,
  experiment_filter_tag = experiment_filter_tag,
  report_stats = TRUE,
  mapping = mapping,
  col_types_map = col_types_map,
  additional_metadata = additional_metadata,
  experiment_spec_cols = experiment_spec_cols,
  load_if_present = TRUE,
  additional_metadata_join_cols =
    c("Metadata_broad_sample"),
  mapping_join_cols =
    c("Metadata_Plate")
) 
```

```{r}
# profiles <-
#   collate_profiles(
#     experiments = experiments,
#     profile_level = "normalized_negcon",
#     batch = params$batch,
#     overwrite_profiles = FALSE,
#     experiment_filter_tag = experiment_filter_tag,
#     report_stats = TRUE,
#     mapping = mapping,
#     col_types_map = col_types_map,
#     additional_metadata = additional_metadata,
#     experiment_spec_cols = experiment_spec_cols,
#     load_if_present = TRUE,
#     additional_metadata_join_cols =
#       c("Metadata_broad_sample"),
#     mapping_join_cols =
#       c("Metadata_Plate")
#   ) 
```

# Load and save Level 4b profiles

```{r}
reprocess_feature_select <-
  function(profile_level_in,
           profile_level_out,
           batch,
           experiment_filter_tag,
           col_types_master,
           overwrite_profiles = FALSE) {
    
    csvgz_all_experiments_file_in <-
      glue(
        "../collated/{batch}/{batch}_{experiment_filter_tag}_{profile_level_in}.csv.gz"
      )
    
    csvgz_all_experiments_file_out <-
      glue(
        "../collated/{batch}/{batch}_{experiment_filter_tag}_{profile_level_out}.csv.gz"
      )
    
    parquet_all_experiments_file_out <-
      glue(
        "../collated/{batch}/{batch}_{experiment_filter_tag}_{profile_level_out}.parquet"
      )
    
    if (!file.exists(parquet_all_experiments_file_out)) {
      if (!file.exists(csvgz_all_experiments_file_out) ||
          overwrite_profiles) {
        futile.logger::flog.info("Run this command")
        futile.logger::flog.info("conda env create --force --file environment.yml")
        futile.logger::flog.info("conda activate pilot-cpjump1-data")
        futile.logger::flog.info(
          glue(
            "python feature_select_pipeline/feature_select_pipeline.py --config feature_select_pipeline/config.yml --input {csvgz_all_experiments_file_in} --output {csvgz_all_experiments_file_out}"
          )
        )
        stop(glue(
          "Run command to generate {csvgz_all_experiments_file_out}"
        ))
        
      } else {
        futile.logger::flog.info(glue("Loading {csvgz_all_experiments_file_out} ..."))
        
        profiles <-
          read_csv(csvgz_all_experiments_file_out,
                   col_types = col_types_master)
        
        futile.logger::flog.info(glue("Writing {parquet_all_experiments_file_out} ..."))
        
        profiles %<>% select(matches("Metadata_"), everything())
        
        profiles %>%
          write_parquet(
            parquet_all_experiments_file_out,
            compression = "gzip",
            compression_level = 9
          )
        
      }
    }
    
    futile.logger::flog.info(glue("Loading {parquet_all_experiments_file_out} ..."))
    
    profiles <-
      read_parquet(parquet_all_experiments_file_out)
    
    report_na_stats(profiles)
    
    n_outlier_variables <-
      length(find_outlier_variables(profiles))
    
    futile.logger::flog.info(
      glue(
        "There are {n_outlier_variables} outlier variables assuming they were z-scored."
      )
    )
  }
```

```{r results='asis'}
profile_level_in <- "normalized"

profile_level_out <- "normalized_feature_select"

profiles <-
  reprocess_feature_select(
    profile_level_in = profile_level_in,
    profile_level_out = profile_level_out,
    batch = "2020_11_04_CPJUMP1",
    experiment_filter_tag = experiment_filter_tag,
    col_types_master = col_types_master
  ) 
```

```{r}
reprocess_outlier_trim <-
  function(profile_level_in,
           profile_level_out,
           experiment_filter_tag,
           batch,
           overwrite_profiles = FALSE) {
  parquet_all_experiments_file_in <-
    glue("../collated/{batch}/{batch}_{experiment_filter_tag}_{profile_level_in}.parquet")
  
  parquet_all_experiments_file_out <-
    glue(
      "../collated/{batch}/{batch}_{experiment_filter_tag}_{profile_level_out}.parquet"
    )
  
  futile.logger::flog.info(glue("Looking for {parquet_all_experiments_file_in}"))
  stopifnot(file.exists(parquet_all_experiments_file_in))
  
  if (!file.exists(parquet_all_experiments_file_out))  {
    futile.logger::flog.info(glue("Loading {parquet_all_experiments_file_in} ..."))
    
    profiles <- read_parquet(parquet_all_experiments_file_in)
    
    outlier_variables <- find_outlier_variables(profiles)
    
    profiles %<>% select(-any_of(outlier_variables))
    
    futile.logger::flog.info(glue("Writing {parquet_all_experiments_file_out} ..."))
    
    profiles %<>% select(matches("Metadata_"), everything())
    
    profiles %>%
      write_parquet(parquet_all_experiments_file_out, compression = "gzip", compression_level = 9)
    
  } else {
    futile.logger::flog.info(glue("Loading {parquet_all_experiments_file_out} ..."))
    
    profiles <-
      read_parquet(parquet_all_experiments_file_out)
    
  }
  
  report_na_stats(profiles)
  
  n_outlier_variables <-
    length(find_outlier_variables(profiles))
  
  futile.logger::flog.info(glue(
    "There are {n_outlier_variables} outlier variables assuming they were z-scored."
  ))
}
```

```{r results='asis'}
profile_level_in  <- "normalized_feature_select"
profile_level_out <- "normalized_feature_select_outlier_trimmed"

profiles <-
  reprocess_outlier_trim(
    profile_level_in = profile_level_in,
    profile_level_out = profile_level_out,
    batch = "2020_11_04",
    experiment_filter_tag = experiment_filter_tag
  ) 
```
